---
title: "HW: Week 4"
author: "36-350 -- Statistical Computing"
date: "Week 4 -- Spring 2021"
output:
  pdf_document:
    toc: no
  html_document:
    theme: spacelab
    toc: yes
    toc_float: yes
---

Name: Jacky Liu

Andrew ID: jackyl1

You must submit **your own** HW as a PDF file on Gradescope.

```{r wrap-hook,echo=FALSE}
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
```

---

## Question 1
*(20 points)*

You are given the following matrix:
```{r linewidth=80}
set.seed(505)
mat = matrix(rnorm(900),30,30)
mat[sample(30,1),sample(30,1)] = NA
```
Compute the standard deviation for each row, using `apply()` and your own on-the-fly function, i.e., a function that is defined *within* the argument list being passed to `apply()`. **Do not use the function sd()!** Realize that since there is a missing value within the matrix, you need to define your function so as to only take into account the non-missing data in each row. If your vector of standard deviations has an `NA` in it, then your function isn't quite working yet.
```{r linewidth=80}
stdev = function(row) {
  # create a new vector with no NA values
  r = row[!is.na(row)]
  # calculate mean
  mean = sum(r) / length(r)
  # implement stdev formula
  summation = 0
  for(num in r){
    summation = summation + (num - mean)^2
  }
  st_dev = sqrt(summation/(length(r)-1))
  return(st_dev)
}
# compute the standard deviation for each row
apply(mat, 1, stdev)
```
---

Below we read in the data on the political economy of strikes.
```{r linewidth=80}
strikes.df = read.csv("http://www.stat.cmu.edu/~mfarag/350/strikes.csv")
```

---

## Question 2
*(20 points)*

Using `split()` and `sapply()`, compute the average unemployment rate, inflation rates, and strike volume for each year represented in the `strikes.df` data frame. The output should be a matrix of dimension 3 $\times$ 35. (You need not display the matrix contents...just capture the output from `sapply()` and pass that output to `dim()`.) Provide appropriate row names (see `rownames()` to your output matrix. Display the columns for 1962, 1972, and 1982. (This can be done in one line as opposed to three.)
```{r linewidth=80}

unemp = strikes.df$unemployment
unemp = split(unemp, f=strikes.df$year)
unemployment.means = sapply(unemp, FUN=mean)
infl = strikes.df$inflation
infl = split(infl, f=strikes.df$year)
inflation.means = sapply(infl, FUN=mean)
strk = strikes.df$strike.volume
strk = split(strk, f=strikes.df$year)
strikevol.means = sapply(strk, FUN=mean)

mat = rbind(unemployment.means, inflation.means, strikevol.means)

mat = data.matrix(mat)
dim(mat)
mat[,c(12,22,32)]
```

## Question 3
*(20 points)*

Utilize piping and `group_by()`, etc., to compute the average unemployment rate for each country, and display that average for only those countries with the maximum and minimum averages. To be clear: your output should only show average unemployment for Ireland and Switzerland, and nothing else. (Hint: remember `slice()`, a less-often-used `dplyr` function.) Hint: arrange your output in order of descending average unemployment, then note that `n()` applied as an argument to the right function will return the last row.
```{r linewidth=80, warning=FALSE}
library(tidyverse)
df = group_by(strikes.df, country) %>% summarize(unemp_rate = mean(unemployment))
df = df[order(-df$unemp_rate),]
df %>% slice(c(1, n()))
```





